Disentangling patterns of gene expression in high grade serous ovarian cancer by matrix factorization

### 2020-08-11 MSc_Dissertation_B156476.tex
Added batch effect results, figs moved to appendix, FloatBarrier, split up gene enrichment figs

\documentclass[tikz, 12pt,a4paper,oneside,fleqn]{article}

@@

\usepackage{caption}
\captionsetup[figure]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\captionsetup[table]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{afterpage}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.4]{Sompairac_ICA_NMF_PCA.png}
\end{center}
\caption{Illustrating the differing results of component extraction by PCA, ICA and NMF.  From \cite{Sompairac2019}}
\label{fig-pca-ica-nmp-explanation}
\end{figure}

@@

\begin{figure}[ht]
\def\M{2.2cm}
\def\N{1.3cm}
\def\K{0.3cm}
\def\fudge{-0.8mm}
\centering
\subfloat[Optimising independence of Meta\underline{genes}]{
\begin{tikzpicture}[node distance = 0.25cm]
	\node (X) [draw, fill=green, minimum height=\M, minimum width=\N] {\large X};
	\node [right=of X] (eq) {\large $\approx$};
	\node [right=of eq] (S) [draw, fill=yellow, minimum width=\K, minimum height=\M] {\large W};
	\node [right=of S] (times) {\LARGE$\times$};
	\node [right=of times] (A) [draw, fill=cyan, minimum width=\N, minimum height=\K] {\large H};

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.7]{methodlology_overview.pdf}
\end{center}
\caption{Overview of methodology as information flow.  The diagramatic convention here is that at each stage, both datasets -- TCGA and AOCS -- are processed separately by the given algorithm.  The exception is Gene Set Intersection, which involves interaction between the datasets.}
\label{fig-method-overview}
\end{figure}

@@

@@

\begin{figure}[ht]
\begin{center}
\begin{Verbatim}[baselinestretch=1, frame=single, rulecolor=\color{blue}, label=Metagene Stability Assessment, fontfamily=courier, fontsize=\small]

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.13]{GO_0042773_lineage.png}
\caption{Example of a small section of the Gene Ontology (GO), focussed GO:0042773 (in purple), showing parents and children of the term.   Generated by GOATOOLS.}
\label{fig-eg_go_graphic}
\end{center}
\end{figure}

@@

\def\S{0.45}
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from ICA facorizations on the N=80 AOCS dataset, comparing \emph{fixed} sampling (top row) with \emph{bootstrap} sampling (bottom row) for a selection of factorization ranks.  }
\label{fig-AOCS-ica-fixed-vs-bootstrap}
\label{fig-first-cluster-result}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_fixed.pdf}} 
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}}
\caption{Clustering of metagenes from ICA facorizations on the N=374 TCGA dataset, again comparing fixed and bootstrap sampling.}
\label{fig-TCGA-ica-fixed-vs-bootstrap}
\end{center}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_NMF_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_PCA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=0.55]{\plotspath{/TCGA_OV_VST/FactorClustering/silhouette_plots_TCGA_2_3_4_5_6_7_8_9_10_bootstrap.pdf}} 
\caption{Metagene clustering for all three methods applied to the N=374 TCGA dataset with bootstrap sampling, over a range of factorization ranks.  The silhouette scores are also plotted (bottom).  It is on the basis of this figure that the factorization ranks for each method were selected.}
\label{fig-TCGA-all-bootstrap}
\label{fig-last-cluster-result}
\end{center}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{TCGA_OV_VST/GeneEnrichment/component_jaccard_heatmap.pdf}}
\end{center}
\caption{Heatmap of Jaccard similarities between the candidate genes identified by the 11 components.}
\label{fig-component-jaccard-similarities}
\end{figure}

@@

Gene enrichment analysis against the GO results in, for each of the 11 metagenes, a table of GO terms with a list of the candidate genes which have an annotated association to that term; this table can be found in the appendix, table \ref{tab-go-enrichment-results}.
To take full advantage of the hierarchical nature of the GO, results are presented as lineage maps in figures \ref{fig-NMF-go-lineage} to \ref{fig-PCA-1-3-go-lineage}.  
The contributing candidate genes for all significantly enriched terms are shown beneath each figure.  
High-level terms at depth less than 3 were removed, since these are generic (e.g. "regulation of biological process") and so uninteresting.   
One component -- ICA-4-of-5 -- was problematic in that 49 GO terms (with depth $\geq 3$) were identified as significant.  
For this component it was necessary to limit the graphic to the 12 terms having the largest number of associated candidate genes.

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.14]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF-1-of-3} \\
\vspace{1cm}
\includegraphics[scale=0.14]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_2_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_2_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF 2-of-3} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components NMF-1-of-3 and NMF-2-of-3.  In these diagrams, enriched terms are coloured purple, while there ancestors in the ontology are yellow.  (NMF-3-of-3 produced no significant enrichment results)}
\label{fig-NMF-go-lineage}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.1]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-3-of-5} \\
\vspace{1cm}
\includegraphics[scale=0.07]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_5_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_5_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-5-of-5} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components ICA-3-of-5 and ICA-5-of-5}
\label{fig-ICA-3-5-go-lineage}
\end{center}
\end{figure}

\begin{sidewaysfigure}[ht]
\begin{center}
\includegraphics[width=\textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_4_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_4_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
\caption{Lineage maps of enriched Gene Ongology (GO) terms for component ICA-4-of-5. For this component it was necessary to additionally filter the enriched terms -- see text.}
\label{fig-ICA-4-go-lineage}
\end{center}
\end{sidewaysfigure}

\begin{figure}[ht]
\begin{center}
\includegraphics[width=0.5 \textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-1-of-3} \\
\vspace{1cm}
\includegraphics[width= \textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-3-of-3} \\
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components 1 and 3 from the rank=5 PCA factorization. (Component PCA-2-of-3 produced no significant enrichment results)  }
\label{fig-PCA-1-3-go-lineage}
\end{center}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_TCGA.pdf}} 
\caption{Kaplan-Meier plots for each metasample component, stratified at the median value, for TCGA $\rightarrow$ TCGA for overall survival (OS) case.  Hazard ratio and p-value is shown for each case.  The final plot (bottom right) is unstratified overall survival}
\label{fig-kaplan-meier-os-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for overall survival (OS).}
\label{fig-kaplan-meier-os-TCGA-AOCS}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_NMF.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_ICA.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_PCA.pdf}}
\caption{Visual summary of survival analysis as applied to TCGA(OS), AOCS(OS) and AOCS(PFS).  Plots are divided by factorization method.  Bar heights show $\log_2(\mbox{HR})$ with p-value also shown.}
\label{fig-summary-survival-plots}
\end{center}
\end{figure}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5,draft=false]{\plotspath{BoxAndScatters/genomic_feature_scatters_TCGA_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples and genomic features.  In the case of WGD which has binary values (0, 1), jitter is applied for visual effect only.   Above each plot is shown a correlation coefficient $r$ and associated p-value, highlighted with '***' where $p < 0.01$.  In the case of WGD , the Point-Biserial correlation is used; for all others Pearson's $r$ is used.}
\label{fig-genomic-scatters}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.7, draft=false]{\plotspath{HeatMaps/clustered_heatmap_TCGA_AOCS.pdf}} 
\caption{Heatmap of the metasamples matrix for the AOCS dataset, computed from metagenes factorized from the TCGA dataset.   Several patient metadata variables are shown in columns to the left.}
\label{fig-heatmap-TCGA-AOCS}
\end{center}
\end{figure}

\begin{table}[ht]
\begin{center}
\begin{tabular}{|l|l|}
\hline
Genomic feature & Significant correlations \\
\hline 
WGD & NMF-1-of-3, ICA-5-of-5, PCA-3-of-3 \\
Cellularity & NMF-1-of-3, ICA-3-of-5, ICA-5-of-5, PCA-3-of-3 \\
HR Detect & NMF-2-of-3, NMF-3-of-3, ICA-2-of-5 \\
Mutational load & None \\
CNV load & None \\
SV load & NMF-2-of-3 \\
\hline
\end{tabular}
\end{center}
\caption{Summary of metasample components showing a significant correlation with each genomic feature.}
\label{tab-genomic-scatter-summary}
\end{table}

@@

\bibliography{OvarianCancer}
\bibliographystyle{ieeetr}

@@

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from NMF facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-nmf-fixed-vs-bootstrap}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_PCA_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_PCA_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from PCA facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-pca-fixed-vs-bootstrap}
\end{center}
\end{figure}

\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_pfs_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for progression free survival (PFS).  }
\label{fig-kaplan-meier-pfs-TCGA-AOCS}
\end{center}
\end{figure}

TODO - I'll likely move the AOCS survivial plots here, and perhaps some of the GO lineages.
\begin{figure}[ht]
\begin{center}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_NMF_TCGA_AOCS.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_ICA_TCGA_AOCS.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_PCA_TCGA_AOCS.pdf}}
\caption{Boxplots showing relationship between metasamples (derived from TCGA metagenes, applied to AOCS data) and whole genome doubling (WGD) status.  Metasamples from NMF, ICA and PCA are shown separately.  To allow convenient visualisation, each metasample is normalised to zero mean unit variance, hence between metasample variation is removed.}
\end{center}
\label{fig-wgd-boxplots}
\end{figure}

>>>>

\documentclass[draft, tikz, 12pt,a4paper,oneside,fleqn]{article}

@@

\usepackage{caption}
\captionsetup[figure]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\captionsetup[table]{labelfont=bf,format=hang,font={stretch=1.2, small}}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{setspace}
\usepackage{placeins}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.4]{Sompairac_ICA_NMF_PCA.png}
\end{center}
\caption{Illustrating the differing results of component extraction by PCA, ICA and NMF.  From \cite{Sompairac2019}}
\label{fig-pca-ica-nmp-explanation}
\end{figure}

@@

\begin{figure}[htb!]
\def\M{2.2cm}
\def\N{1.3cm}
\def\K{0.3cm}
\def\fudge{-0.8mm}
\centering
\subfloat[Optimising independence of Meta\underline{genes}]{
\begin{tikzpicture}[node distance = 0.25cm]
	\node (X) [draw, fill=green, minimum height=\M, minimum width=\N] {\large X};
	\node [right=of X] (eq) {\large $\approx$};
	\node [right=of eq] (S) [draw, fill=yellow, minimum width=\K, minimum height=\M] {\large W};
	\node [right=of S] (times) {\LARGE$\times$};
	\node [right=of times] (A) [draw, fill=cyan, minimum width=\N, minimum height=\K] {\large H};

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.7]{methodlology_overview.pdf}
\end{center}
\caption{Overview of methodology as information flow.  The diagramatic convention here is that at each stage, both datasets -- TCGA and AOCS -- are processed separately by the given algorithm.  The exception is Gene Set Intersection, which involves interaction between the datasets.}
\label{fig-method-overview}
\end{figure}

@@

\FloatBarrier

@@

\begin{figure}[htb!]
\begin{center}
\begin{Verbatim}[baselinestretch=1, frame=single, rulecolor=\color{blue}, label=Metagene Stability Assessment, fontfamily=courier, fontsize=\small]

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.13]{GO_0042773_lineage.png}
\caption{Example of a small section of the Gene Ontology (GO), focussed GO:0042773 (in purple), showing parents and children of the term.   Generated by GOATOOLS.}
\label{fig-eg_go_graphic}
\end{center}
\end{figure}

@@

\def\S{0.45}
\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_ICA_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from ICA facorizations on the N=80 AOCS dataset, comparing \emph{fixed} sampling (top row) with \emph{bootstrap} sampling (bottom row) for a selection of factorization ranks.  }
\label{fig-AOCS-ica-fixed-vs-bootstrap}
\label{fig-first-cluster-result}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_fixed.pdf}} 
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}}
\caption{Clustering of metagenes from ICA facorizations on the N=374 TCGA dataset, again comparing fixed and bootstrap sampling.}
\label{fig-TCGA-ica-fixed-vs-bootstrap}
\end{center}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_NMF_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_ICA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=\S]{\plotspath{TCGA_OV_VST/FactorClustering/multiple_single_factors_scatter_TCGA_PCA_2_3_5_10_bootstrap.pdf}} \\
\includegraphics[scale=0.55]{\plotspath{/TCGA_OV_VST/FactorClustering/silhouette_plots_TCGA_2_3_4_5_6_7_8_9_10_bootstrap.pdf}} 
\caption{Metagene clustering for all three methods applied to the N=374 TCGA dataset with bootstrap sampling, over a range of factorization ranks.  The silhouette scores are also plotted (bottom).  It is on the basis of this figure that the factorization ranks for each method were selected.}
\label{fig-TCGA-all-bootstrap}
\label{fig-last-cluster-result}
\end{center}
\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{TCGA_OV_VST/GeneEnrichment/component_jaccard_heatmap.pdf}}
\end{center}
\caption{Heatmap of Jaccard similarities between the candidate genes identified by the 11 components.}
\label{fig-component-jaccard-similarities}
\end{figure}

\FloatBarrier

@@

Gene enrichment analysis against the GO results in, for each of the 11 metagenes, a table of GO terms with a list of the candidate genes which have an annotated association to that term; this table can be found in the appendix, table \ref{tab-go-enrichment-results}.
To take full advantage of the hierarchical nature of the GO, results are presented as lineage maps in figures \ref{fig-NMF_1-go-lineage} to \ref{fig-PCA-3-3-go-lineage}.  
The contributing candidate genes for all significantly enriched terms are shown beneath each figure.  
High-level terms at depth less than 3 were removed, since these are generic (e.g. "regulation of biological process") and so uninteresting.   
One component -- ICA-4-of-5 -- was problematic in that 49 GO terms (with depth $\geq 3$) were identified as significant.  
For this component it was necessary to limit the graphic to the 12 terms having the largest number of associated candidate genes.

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.1]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF-1-of-3} \\
\end{center}
\caption{Lineage maps of enriched Gene Ongology (GO) terms for components NMF-1-of-3  In these diagrams, enriched terms are coloured purple, while there ancestors in the ontology are yellow.  }
\label{fig-NMF_1-go-lineage}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.1]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_2_of_NMF_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_2_of_NMF_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component NMF 2-of-3} \\
\caption{Lineage maps of enriched GO terms for components NMF-2-of-3.   (NMF-3-of-3 produced no significant enrichment results)}
\label{fig-NMF_2-go-lineage}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.1]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-3-of-5} \\
\end{center}
\caption{Lineage maps of enriched GO terms for components ICA-3-of-5}
\label{fig-ICA-3-5-go-lineage}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.07]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_5_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_5_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component ICA-5-of-5} \\
\caption{Lineage maps of enriched GO terms for components ICA-5-of-5}
\label{fig-ICA-5-5-go-lineage}
\end{center}
\end{figure}

\begin{sidewaysfigure}[htb!]
\begin{center}
\includegraphics[width=\textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_4_of_ICA_5.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_4_of_ICA_5.txt}}
\end{spacing}
\vspace{0.5cm}
\caption{Lineage maps of enriched GO terms for component ICA-4-of-5. For this component it was necessary to additionally filter the enriched terms -- see text.}
\label{fig-ICA-4-go-lineage}
\end{center}
\end{sidewaysfigure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[width=0.4 \textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_1_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_1_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-1-of-3} \\
\end{center}
\caption{Lineage maps of enriched GO terms for component PCA-1-of-3  }
\label{fig-PCA-1-3-go-lineage}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[width= \textwidth]{\plotspath{TCGA_OV_VST/GeneEnrichment/go_lineage_3_of_PCA_3.png}} \\
\begin{spacing}{0.6}
{\small Genes: \tiny \input{/home/ipoole/Documents/gitrepos/HgsocTromics/Plots/TCGA_OV_VST/GeneEnrichment/go_gene_set_3_of_PCA_3.txt}}
\end{spacing}
\vspace{0.5cm}
{\small Component PCA-3-of-3} \\
\end{center}
\caption{Lineage maps of enriched GO terms for component PCA-3-of-3. (Component PCA-2-of-3 produced no significant enrichment results)  }
\label{fig-PCA-3-3-go-lineage}

\end{figure}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for overall survival (OS).}
\label{fig-kaplan-meier-os-TCGA-AOCS}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_NMF.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_ICA.pdf}} \\
\includegraphics[scale=0.5]{\plotspath{/SurvivalAnalysis/survival_summary_barplots_PCA.pdf}}
\caption{Visual summary of survival analysis as applied to TCGA(OS), AOCS(OS) and AOCS(PFS).  Plots are divided by factorization method.  Bar heights show $\log_2(\mbox{HR})$ with p-value also shown.}
\label{fig-summary-survival-plots}
\end{center}
\end{figure}

\FloatBarrier

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5,draft=false]{\plotspath{BoxAndScatters/genomic_feature_scatters_TCGA_AOCS.pdf}} 
\caption{Grid of scatter plots to visualise the correlation between metasamples and genomic features.  In the case of WGD which has binary values (0, 1), jitter is applied for visual effect only.   Above each plot is shown a correlation coefficient $r$ and associated p-value, highlighted with '***' where $p < 0.01$.  In the case of WGD , the Point-Biserial correlation is used; for all others Pearson's $r$ is used.}
\label{fig-genomic-scatters}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.7, draft=false]{\plotspath{HeatMaps/clustered_heatmap_TCGA_AOCS.pdf}} 
\caption{Heatmap of the metasamples matrix for the AOCS dataset, computed from metagenes factorized from the TCGA dataset.   Several patient metadata variables are shown in columns to the left.}
\label{fig-heatmap-TCGA-AOCS}
\end{center}
\end{figure}

\begin{table}[htb!]
\begin{center}
\begin{tabular}{|l|l|}
\hline
Genomic feature & Significant correlations \\
\hline 
WGD & NMF-1-of-3, ICA-5-of-5, PCA-3-of-3 \\
Cellularity & NMF-1-of-3, ICA-3-of-5, ICA-5-of-5, PCA-3-of-3 \\
HR Detect & NMF-2-of-3, NMF-3-of-3, ICA-2-of-5 \\
Mutational load & None \\
CNV load & None \\
SV load & NMF-2-of-3 \\
\hline
\end{tabular}
\end{center}
\caption{Summary of metasample components showing a significant correlation with each genomic feature.}
\label{tab-genomic-scatter-summary}
\end{table}

\FloatBarrier

\section{Matrix factorization discovers and effectively removes batch effects}

Here are presented results of applying the three MF methods to the combined AOCS and TCGA datasets.

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.7, draft=false]{\plotspath{HeatMaps/clustered_heatmap_BOTH_BOTH.pdf}} 
\caption{Heatmap of the metasamples matrix for the combined AOCS + TCGA dataset.  The 'Which' field indicates the source dataset; purple: AOCS, yellow:TCGA.}
\label{fig-heatmap-BOTH}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_NMF_BOTH_BOTH.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_ICA_BOTH_BOTH.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_PCA_BOTH_BOTH.pdf}}
\caption{Boxplots of metasamples from the combined AOCS + TCGA datasets, showing relationship between metasamples dataset origin; Which=0 or 1 for AOCS or TCGA respectively.}
\end{center}
\label{fig-wgd-boxplots-BOTH}
\end{figure}

From the clustered heatmap of figure \ref{fig-heatmap-BOTH} it is clear that the five components clustered to the left are highly sensitive to the batch effect.

@@

\small
\bibliography{OvarianCancer}
\bibliographystyle{ieeetr}

@@

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_NMF_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from NMF facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-nmf-fixed-vs-bootstrap}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_PCA_2_3_5_10_fixed.pdf}} \\
\includegraphics[scale=\S]{\plotspath{AOCS_Protein/FactorClustering/multiple_single_factors_scatter_AOCS_PCA_2_3_5_10_bootstrap.pdf}}\\
\caption{Clustering of metagenes from PCA facorizations on the AOCS dataset, comparing fixed and bootstrap sampling}
\label{fig-AOCS-pca-fixed-vs-bootstrap}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_os_TCGA_TCGA.pdf}} 
\caption{Kaplan-Meier plots for each metasample component, stratified at the median value, for TCGA $\rightarrow$ TCGA for overall survival (OS) case.  Hazard ratio and p-value is shown for each case.  The final plot (bottom right) is unstratified overall survival}
\label{fig-kaplan-meier-os-TCGA-TCGA}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.5]{\plotspath{SurvivalAnalysis/multiple_kaplan_meier_pfs_TCGA_AOCS.pdf}} 
\caption{Kaplan-Meier plots for TCGA $\rightarrow$ TCGA for progression free survival (PFS).  }
\label{fig-kaplan-meier-pfs-TCGA-AOCS}
\end{center}
\end{figure}

\begin{figure}[htb!]
\begin{center}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_NMF_TCGA_AOCS.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_ICA_TCGA_AOCS.pdf}}
\includegraphics[scale=0.6]{\plotspath{BoxAndScatters/wgd_boxplot_PCA_TCGA_AOCS.pdf}}
\caption{Boxplots showing relationship between metasamples (derived from TCGA metagenes, applied to AOCS data) and whole genome doubling (WGD) status.  Metasamples from NMF, ICA and PCA are shown separately.  To allow convenient visualisation, each metasample is normalised to zero mean unit variance, hence between metasample variation is removed.}
\end{center}
\label{fig-wgd-boxplots}
\end{figure}
